Efficient, Robust and Exact Boolean Operations on N Triangular Mesh Primitives

### 2016-06-12 summary.tex
before delete equation 3.

As we discussed in Section \ref{sec:paradigm}, one of the problem in Boolean evaluation under vertex-based representation is it inevitably introduces new vertices, whose coordinates cannot be exactly represented using standard float-point number. On the other hand, if plane-based representation is used, no new geometry information needs to be constructed---the result mesh is composed of the planes of the input primitives. Therefore, we can use only geometric decision predicates to perform the Boolean operation.

Using plane-based representation, each face of polyhedron is represented by a supporting plane $S$, where the face lies, and a list bounding planes $\{B_i\}$, each corresponding an edge of the face. Edge line is represented by intersection of $S$ and the related $B_i$. Corner point is then the intersection of $S$ and two consecutive bounding planes $B_i$ and $B_{i+1}$. Efficient and exact conversion from the vertex-based representation to plane-based representation is described in detail in \cite{campen2010exact}. If the input vertex coordinates precision is $L$, the maximum precision $P$ required to represent the plane coefficients is:
\begin{equation}
\label{eq:precision}
P = (L-1)+2(K-1)+3+1,
\end{equation}
where $K$  is the maximum bits to represent edge vector coordinates. If the plane coefficient has $M$ bit precision, $M$ have to greater than $P$ for exact representation. We rigorously enforce the criterion to $M \ge P+1$, and we will explain it in section \ref{sec:embed}.

Let $\delta = 2^{K-L-1}$ be the relative length (in max-norm) of the
longest edge in the mesh (relative to the bounding box). If we use IEEE 754 double precision ($M=53$) and assume $\delta \le 2^{-5.5}$, the maximum input precision $L$ can reach to 20, which is enough for most CAD application. As the precision of input coordinates is finite, we can determine in advance the upper boundary of precision needed to perform a precise decision, with which we can use fixed precision predicates techniques to gain much performance while keeping exactness. We implement the predicates using filtering techniques proposed by Shewchuk \cite{shewchuk1997adaptive}.

\subsection{Substrates}
\label{sec:substrates}
Our method rigorously follows the paradigm of plane-based geometry. All decision predicates are computed using planes as input. These predicates include checking coincidence and co-orientation of planes, orientation of a plane with respect to a point, and whether three planes intersect in a unique point, etc. We omit the discussion of the predicates which is already in prior work \cite{bernstein2009fast,banerjee1996topologically}, and focus on specific predicates in our method in the following.

Before we start the discussion, we first introduce several regulations frequently used in the rest of our paper. A plane $P_x$ can be represented by plane coefficients $(a_x, b_x, c_x, d_x)$. The normal of $P_x$ is denoted as $\vec{N}_x$. Under plane-based representation, a line $L$ can be represented by intersection of two planes $(P_m, P_n)$, or in short $L: (P_m, P_n)$. The positive direction of $L$ is defined by cross product of the normals of $P_a$ and $P_b$, that is, $\vec{L}=(a_m, b_m, c_m) \times (a_n, b_n, c_n)$. A point $p$ is represented by plane triples $(P_a, P_b, P_c)$, or in short $p: (P_a, P_b, P_c)$.

Given two points $p$ and $q$ in the same edge line $L: (P_u, P_v)$, the two points can be represented by plane triples $(P_u, P_v, P_p)$ and $(P_u, P_v, P_q)$, sharing the same first two components. We want to determine if $q$ is in positive side of $p$ along the $\vec{L}$. We call this two-point comparison. This predicates can be converted to the problem of orientation of a plane $P_p$ with respect to a point $q$, except that there is an extra request---we have to guarantee cosine between $\vec{L}$ and the normal $\vec{N}_p$ be positive, that is:
\begin{equation}
\label{eq:pointcomp}
   \left|
  \begin{array}{ccc}   %该矩阵一共3列，每一列都居中放置
  a_u  & b_u & c_u \\
  a_v  & b_v & c_v \\
  a_p  & b_p & c_p
  \end{array}
  \right| > 0,
\end{equation}
If not, we negate $P_p$ to be the test plane. The computation of this predicates is described in \cite{bernstein2009fast}.

>>>>

As we discussed in Section \ref{sec:paradigm}, one of the problem in Boolean evaluation under vertex-based representation is it inevitably introduces new vertices, whose coordinates cannot be exactly represented using common float-point number. On the other hand, if plane-based representation is used, no new geometry information needs to be constructed. Therefore, we can use only geometric decision predicates to perform the boundary evaluation, which means it is exact and robust.

Using plane-based representation, each face of polyhedron is represented by a supporting plane $p^s$, where the face lies, and a list bounding planes $\{p^b_i \ \vert\  i = 1, 2, 3\}$, each has  a corresponding edge. Edge line is represented by intersection of $p^s$ and the related $p^b_i$. Corner vertex is then the intersection of $p^s$ and two consecutive bounding planes $p^b_i$ and $p^b_{i+1}$. Efficient and exact conversion from the vertex-based representation to plane-based representation is described in detail in \cite{campen2010exact}. If the input vertices precision is $L$, the maximum precision $P$ required to represent the plane coefficients is:
\begin{equation}
\label{eq:precision}
P = (L-1)+2(K-1)+3+1,
\end{equation}
where $K$  is the maximum bits to represent edge vector components. If the plane coefficient has $M$ bit precision, $M$ have to greater than or equal to $P$ for exact conversion. However, different from \cite{campen2010exact}, we choose to enforce the criterion to $M \ge P+1$, and we will explain the enforcement in section \ref{sec:embed}.

Let $\delta = 2^{K-L-1}$ be the relative length (in max-norm) of the
longest edge in the mesh (relative to the bounding box). If we use IEEE 754 double precision ($M=53$) and assume $\delta \le 2^{-5.5} \approx 0.22$, the maximum input precision $L$ can reach to 20, which is enough for most application. As the precision of input coordinates is finite, we can determine in advance the upper boundary of precision needed to perform a precise decision, with which we can use fixed precision predicates techniques to gain much performance while keeping exactness. We implement the predicates using filtering techniques proposed by Shewchuk \cite{shewchuk1997adaptive}.

\subsection{Substrates}
\label{sec:substrates}
Our method rigorously follows the paradigm of plane-based geometry. Predicates are all computed using planes as input. These predicates include checking coincidence and co-orientation of planes, orientation of a plane with respect to a point, and whether three planes intersect in a unique point, etc. In the following, we only focus on two specific predicates. Other substrates may refer to \cite{bernstein2009fast}.

Before we start the discussion, we first introduce several regulations frequently used in the rest of our paper. A plane $p_x$ can be represented by plane coefficients $(a_x, b_x, c_x, d_x)$. The normal of $p_x$ is denoted as $\vec{n}_x$. Under plane-based representation, a line $l$ can be represented by intersection of two planes $(p_m, p_n)$, or in short, $l: (p_m, p_n)$. The positive direction of $l$ is defined by cross product of the normals of $p_a$ and $p_b$, that is, $\vec{l} = \vec{n}_a \times \vec{n}_b$. A point $v$ is represented by plane triples $(p_a, p_b, p_c)$, or in short, $v: (p_a, p_b, p_c)$.

Suppose there is an edge $e_i(f)$ of a face $f$, and the line of $e_i(f)$, denoted as $l_{i}(f)$, can represented as intersection of supporting plane $p^s(f)$ and bounding plane $p_i^b(f)$. Given two arbitrary points $v_p$ and $v_q$ on $l_{i}(f)$, they can be represented by plane triples $(p^s(f), p_i^b(f), p_p)$ and $(p^s(f), p_i^b(f), p_q)$, sharing the same first two components. We want to determine if $v_q$ is in positive side of $v_p$ along $\vec{l}_{i}(f)$. We call this predicate as \emph{two-point-one-line comparison}. This problem can be converted to the problem of orientation of a plane $p_p$ with respect to a point $v_q$, except that there is an extra request---we have to guarantee cosine between $\vec{l}_{i}(f)$ and the $\vec{n}_p$ (normal of plane $p_p$) be positive, that is:
\begin{equation}
\label{eq:pointcomp}
   \left|
  \begin{array}{ccc}   %该矩阵一共3列，每一列都居中放置
  a_u  & b_u & c_u \\
  a_v  & b_v & c_v \\
  a_p  & b_p & c_p
  \end{array}
  \right| > 0,
\end{equation}
If not, we negate $P_p$ to be the test plane.
